---
title: "FLOW"
author: "Francesco Cabras"
date: "10/19/2020"
output: html_document
---

```{r setup, include=FALSE}
library(WDI)
```

```{r}
W <- WDI(indicator = c("IC.ELC.PRI.KH.DB1619", "EG.IMP.CONS.ZS", "EN.ATM.CO2E.KD.GD", "EG.USE.ELEC.KH.PC", "NY.GDP.PCAP.PP.CD", "EP.PMP.SGAS.CD", "EG.ELC.PETR.ZS", "EG.ELC.RNWX.KH", "EG.ELC.COAL.ZS"), start = 2014, end = 2019, extra = TRUE) 

names(W) <- c("iso2c", "country", "year", "PRI", "DEP", "EMI","CON", "GDP", "OILp", "OILs", "REN", "COA", "iso3c", "region", "capital", "longitude", "latitude", "income", "lending")
names(W)
```

Let us remove cities and aggregate regions from the data set.

```{r}
W <- W[W$income != "Aggregates",]
W <- W[!is.na(W$income),]
```

Filling dependency NAs. If NA is found substitute with latest value from same country. If no data regarding country is found sub with average of the region in consideration.

```{R}
for (row in 1:nrow(W)) {
  area = W$region[row]
  if (W$year[row] == 2014 & is.na(W$DEP[row])) {
    W$DEP[row] <- mean(W[W$region == area,"DEP"],na.rm=T)
  }
}

for (row in 1:nrow(W)) {
  country = W$country[row]
  if (is.na(W$DEP[row])) {
    if (is.na(W[W$country == country & W$year == 2015,"DEP"])) {
      W$DEP[row] <- W[W$country == country & W$year == 2014,"DEP"]
    }
    else {
      W$DEP[row] <- W[W$country == country & W$year == 2015,"DEP"]
    }
  }
} 
```

Filling consumption NAs. If NA is found substitute with latest value from same country. If no data regarding country is found sub with average of the as wealthy countries in consideration.

```{R}
for (row in 1:nrow(W)) {
  area = W$income[row]
  if (W$year[row] == 2014 & is.na(W$CON[row])) {
    W$CON[row] <- mean(W[W$income == area,"CON"],na.rm=T)
  }
}

for (row in 1:nrow(W)) {
  country = W$country[row]
  if (is.na(W$CON[row])) {
    if (is.na(W[W$country == country & W$year == 2015,"CON"])) {
    W$CON[row] <- W[W$country == country & W$year == 2014,"CON"]
    }
    else {
      W$CON[row] <- W[W$country == country & W$year == 2015,"CON"]
    }
  }
} 
```

Filling oil, ren and coal shares  NAs. If NA is found substitute with latest value from same country. If no data regarding country is found sub with average of the region in consideration.

```{R}
for (row in 1:nrow(W)) {
  area = W$region[row]
  if (W$year[row] == 2014 & is.na(W$OILs[row])) {
    W$OILs[row] <- mean(W[W$region == area,"OILs"],na.rm=T)
  }
  if (W$year[row] == 2014 & is.na(W$REN[row])) {
    W$REN[row] <- mean(W[W$region == area,"REN"],na.rm=T)
  }
  if (W$year[row] == 2014 & is.na(W$COA[row])) {
    W$COA[row] <- mean(W[W$region == area,"COA"],na.rm=T)
  }
}

for (row in 1:nrow(W)) {
  country = W$country[row]
  if (is.na(W$OILs[row])) {
    if (is.na(W[W$country == country & W$year == 2015,"OILs"])) {
    W$OILs[row] <- W[W$country == country & W$year == 2014,"OILs"]
    }
    else {
      W$OILs[row] <- W[W$country == country & W$year == 2015,"OILs"]
    }
  }
  if (is.na(W$REN[row])) {
    if (is.na(W[W$country == country & W$year == 2015,"REN"])) {
    W$REN[row] <- W[W$country == country & W$year == 2014,"REN"]
    }
    else {
      W$REN[row] <- W[W$country == country & W$year == 2015,"REN"]
    }
  }
  if (is.na(W$COA[row])) {
    if (is.na(W[W$country == country & W$year == 2015,"COA"])) {
    W$COA[row] <- W[W$country == country & W$year == 2014,"COA"]
    }
    else {
      W$COA[row] <- W[W$country == country & W$year == 2015,"COA"]
    }
  }
} 
```

Filling oil price NAs. If NA is found substitute with latest value from same country. If no data regarding country is found sub with average of the as wealthy countries in consideration.

```{R}
for (row in 1:nrow(W)) {
  area = W$income[row]
  country = W$country[row]
  if (W$year[row] == 2014 & is.na(W$OILp[row])) {
    if (is.na(W[W$country == country & W$year == 2016,"OILp"])) {
    W$OILp[row] <- mean(W[W$income == area,"OILp"],na.rm=T)
    } else {
      W$OILp[row] <- W[W$country == country & W$year == 2016,"OILp"]
    }
  }
}

for (row in 1:nrow(W)) {
  country = W$country[row]
  if (is.na(W$OILp[row])) {
    if (is.na(W[W$country == country & W$year == 2016,"OILp"])) {
    W$OILp[row] <- W[W$country == country & W$year == 2014,"OILp"]
    }
    else {
      W$OILp[row] <- W[W$country == country & W$year == 2016,"OILp"]
    }
  }
} 
```

```{r}
W <- W[!is.na(W$PRI),]
W <- W[!is.na(W$GDP),]
```

Filling emission intensity NAs. If NA is found substitute with latest value from same country. If no data regarding country is found sub with average of countries in that region.

```{R}
for (row in 1:nrow(W)) {
  area = W$region[row]
  if (W$year[row] == 2016 & is.na(W$EMI[row])) {
    W$EMI[row] <- mean(W[W$region == area,"EMI"],na.rm=T)
  }
}

for (row in 1:nrow(W)) {
  country = W$country[row]
  if (is.na(W$EMI[row])) {
    W$EMI[row] <- W[W$country == country & W$year == 2016,"EMI"]
  }
}

```

Write CSV.

```{r}
sum(is.na(W))
write.csv(W, "DAT.csv", row.names=F)
```